### 2024-01-19 paper.tex

\subsection{Analysis of scBS datasets for benchmarks}
To analyze scBS data from \citet{kremer_scnmt}, single-cell CpG methylation reports from all conditions were first stored with \texttt{scbs prepare} and then smoothed with \texttt{scbs smooth} using the default bandwidth of 1000~bp.
\new{
Methylation was then quantified in four sets of genomic intervals:
VMRs, obtained with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$, minimum cell requirement $=6$);
non-overlapping tiles of 100~kb width;
promoter regions as defined by the \textpm2~kb domain around the TSS of coding genes;
and candidate cis-regulatory regions annotated by the ENCODE consortium \citep{encode2020expanded},
Averaged m}ethylation fractions and the shrunken means of residuals were quantified with \texttt{scbs matrix} \new{for all four sets of regions}.
The same analysis was repeated for \new{three additional scBS data sets \citep{luo2017single, bian2018single, argelaguet2019gastru}}, and for smaller data sets generated by randomly sub-sampling cells separately from \new{these} data sets.
\todo{fix: For each data set, we thus obtained four methylation matrices:
either 100~kb tiles or VMRs, quantified either by simple averaging (methylation fractions) or with the shrunken mean of the residuals.
All four matrices were then subjected to PCA.
In the case of methylation fractions, we closely followed methods described in \citet{luo2017single}:
We selected tiles spanning $\ge20$ CpG sites and with sequencing coverage in at least 70\% of cells.
We then imputed missing values by imputing the mean for each tile, centered the values, and performed PCA.
In the case of VMRs, we used iterative PCA to handle missing values, as described earlier.}
For visualization, the first 15 PCs were subjected to UMAP with parameters min\_dist $=0.2$, init $=$ ``spca''.
To flexibly adapt to data sets of different sizes, we set n\_neighbors  $=\frac{\sqrt{n}}{1.5}$ (rounded to the nearest integer), where $n$ is the total cell number.

@@

\noindent%
\begin{minipage}{\linewidth}
	\begin{center}
		\includegraphics[width=.85\textwidth]{figures/SFig_100k_and_CH.png}
	\end{center}
	\captionof{figure}{\small \new{\textbf{\texttt{scbs} performs well on CH-methylation data and on extremely large data sets.}\\
    \textbf{(A)} UMAPs obtained when analysing data from \citet{luo2017single} with three different strategies for producing a methylation matrix:
	top: averaging CH-methylation (mCH) in 100~kb genomic tiles \citep[as done in][to obtain the depicted cell type labels]{luo2017single};
	center: using the shrunken mean of residuals to quantify mCH in mCH-VMRs;
	bottom: shrunken mean of residuals to quantify CpG methylation in CpG-VMRs.
	\textbf{(B)} UMAP obtained when analysing 100\,350 neural single-cell methylomes \citep{liu2021dna} with the default \texttt{scbs} workflow.
	The depicted cell labels were reported by \citet{liu2021dna}.
	\todo{Analysing this large data set took approximately XXX hours and required approximately XXX GB of RAM. (Sveta, pls give some rough estimate here)}
	}}
	\label{fig:umaps}
\end{minipage}

>>>>

\subsection{Analysis of scBS datasets for benchmarks}
To analyze scBS data from \citet{kremer_scnmt}, single-cell CpG methylation reports from all conditions were first stored with \texttt{scbs prepare} and then smoothed with \texttt{scbs smooth} using the default bandwidth of 1000~bp.
\new{
This data was then analysed multiple times with different combinations of analysis methods, namely four ways to divide the genome into intervals, two ways to quantify methylation in these intervals, and four approaches for dimensionality reductions.
The following four sets of genomic intervals were used:
VMRs, obtained with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$, minimum cell requirement $=6$);
adjacent tiles of 100~kb width;
promoter regions as defined by the \textpm2~kb domain around the TSS of coding genes;
and candidate cis-regulatory regions annotated by the ENCODE consortium \citep{encode2020expanded}.
Methylation was quantified either by averaging binary methylation values, or by calculating the shrunken mean of the residuals as described earlier.
}
\todo{...describe the four dimensionality reduction approaches, see below...}
The same analysis was repeated for \new{three additional scBS data sets \citep{luo2017single, bian2018single, argelaguet2019gastru}}, and for smaller data sets generated by randomly sub-sampling cells separately from \new{these} data sets.
\todo{update this and maybe move it up?: For each data set, we thus obtained four methylation matrices:
either 100~kb tiles or VMRs, quantified either by simple averaging (methylation fractions) or with the shrunken mean of the residuals.
All four matrices were then subjected to PCA.
In the case of methylation fractions, we closely followed methods described in \citet{luo2017single}:
We selected tiles spanning $\ge20$ CpG sites and with sequencing coverage in at least 70\% of cells.
We then imputed missing values by imputing the mean for each tile, centered the values, and performed PCA.
In the case of VMRs, we used iterative PCA to handle missing values, as described earlier.
For visualization, the first 15 PCs were subjected to UMAP with parameters min\_dist $=0.2$, init $=$ ``spca''.
To flexibly adapt to data sets of different sizes, we set n\_neighbors  $=\frac{\sqrt{n}}{1.5}$ (rounded to the nearest integer), where $n$ is the total cell number.
}

@@

\noindent%
\begin{minipage}{\linewidth}
	\begin{center}
		\includegraphics[width=.85\textwidth]{figures/SFig_100k_and_CH.png}
	\end{center}
	\captionof{figure}{\small \new{\textbf{\texttt{scbs} performs well on CH-methylation data and on extremely large data sets.}\\
    \textbf{(A)} UMAPs obtained when analysing data from \citet{luo2017single} with three different strategies for producing a methylation matrix:
	top: averaging CH-methylation (mCH) in 100~kb genomic tiles \citep[as done in][to obtain the depicted cell type labels]{luo2017single};
	center: using the shrunken mean of residuals to quantify mCH in mCH-VMRs;
	bottom: shrunken mean of residuals to quantify CpG methylation in CpG-VMRs.
	\textbf{(B)} UMAP obtained when analysing 100\,350 neural single-cell methylomes \citep{liu2021dna} with the default \texttt{scbs} workflow.
	The depicted cell labels were reported by \citet{liu2021dna}.
	\todo{Analysing this large data set took approximately two days on a cluster with 256 GB RAM and and 48 CPUs.}
	}}
	\label{fig:umaps}
\end{minipage}
